### 2021-11-10 documentos/pce.tex
Comentario para Carlos
starting the service is always better

The construction of quantum channels includes basically three ingredients:
linearity, trace preserving and complete positivity. The first one is needed to
send convex combinations of density matrices into the convex combination of the
evolution of such density matrices, this is, let $\mcE$ such linear operation,
thus $p \rho_1 +(1-p) \rho_2 \mapsto p \mcE[\rho_1]+(1-p)\mcE[\rho_2]$. The
trace preserving is required for the process $\mcE$ to be deterministic, this
is, $\tr\mcE[\rho]=\tr \rho =1$ is the probability that the process $\mcE$ will
happen to the state $\rho$, with $\rho \in \mcS(\hilbert)$. The complete
positivity condition is needed to preserve positive semidefiniteness and handle
the non-local character of quantum theory. This is, with positivity of the map
$\mcE$ one has $\mcE[\Delta]\geq 0 \ \ \forall \Delta \in \mcB(\hilbert)$ with
$\Delta\geq 0$. In the other hand, complete positvity of $\mcE$ reads $\id_k
\otimes \mcE[\omega]\geq 0$ for $k=1,2,\dots$, where $\omega$ is the Bell state
between the system and an ancilla with Hilbert space $\hilbert_k$. Therefore,
complete positivity ensures that every positive semidefinite operator of any
extension of the system (where the quantum channel is now a local operation) is
mapped into positive semidefinite operators, trace preserving is automatically
fulfilled for all extensions if $\mcE$ is. Thus, any density matrix of a system
where $\mcE$ acts locally, is sent to a valid density matrix, this is the
substance of complete positivity.

Quantum channels can be characterized by means of the Choi-\jami{} isomorphism~\cite{choi,jamil}. This is, given a linear operation $\mcE:\mcB(\hilbert)\to \mcB(\hilbert)$, the so-called Choi matrix of it, $C_\mcE=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$, where $\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is the Bell state between to copies of \hilbert, is positive semi-definite if and only if $\mcE$ is completely positive.

Quantum channels have the well-known \textit{Kraus representation} (or operator-sum representation) which reads $\mcE[\rho]=\sum_i K_i \rho K^\dagger{}_i$ in fact, a linear operation is completely positive if and only if it has a Kraus representation [cite Kraus book]

@@

\begin{itemize}
\item Canales de Ruskai Pauli maps, que la interseccion es no trivial (no estan contenidos en ninguna direccion, pero la interseccion es no vacia)
y Relación con canales citados en \cite{Siudzinska2020}? \cpnote{Esto quiza lo tengamos
que mover si es que necesitamos herramientas que discutiremos mas adelante para
hacer notar la diferencia}
\end{itemize}

@@

%  $r_{\vec \alpha} =  $
% where $r_{0,\ldots,0}=1$ ($\Tr\rho=1$) and $\sigma_0=\openone$.

Consider the linear operations that simply attenuate the Pauli components of 
a multiqubit system:
\begin{equation}
	\paulicomponents \mapsto \taus \paulicomponents.
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
\label{eq:definition:Pauli}
\end{equation}
These operations are diagonal in the basis $\left\{ 2^{-N/2}\sigma_{\vec \alpha} \right\}$, this follows straightforwardly from \eref{eq:sigma_property}, in fact, the diagonal components are exactly $\taus$, this is $\hat \mcE=\diag(\taus)$. Additionally, they are unital (leave invariant the complete mixture), in fact, as we will show later, they belong to the subset of unital channels that can be written as convex combinations of unitary channels.
 \cpnote{davalos, dale aca un pequeño rollo relevante quiza de una o dos
frases. También debemos tener un nombre pues lo de la diagonazliación funciona
para eso y quizá también seran relevantes para la parte de decoherencia}
\ddnote{De momento no he encontrado alguna reference, para qubits si, pero para
los generales no, estoy checando de nuevo el paper de Ruskai para entender que
tanto se solapan, a ver si puedo decir algo}.
For the operation to preserve the trace of the density matrix, we require that 
$\tau_{\vec 0}=1$. 
A Pauli component erasing (PCE) operation is a Pauli diagonal operation
%  A linear operation 
% $\mcE$ 
that either preserves or completely erases the Pauli components of a
density matrix. That is, 
% it transforms its Pauli components as 
\begin{equation} \label{eq:PCE_definition}
% \begin{split}
% 	\paulicomponents \mapsto \ \taus \paulicomponents,\hspace{.5cm}\\
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
	\paulicomponents \mapsto \taus \paulicomponents,\quad 
	\taus = 0,1.
% \end{split}
\end{equation}
It is worth noticing that not all PCE maps are quantum operations. For single 
qubits, if just one of the $\taus=0$, the complete positivity condition 
is not met, as discussed in \sref{sec:single:qubit}. It will be of central interest
in this article to determine which $\taus$ lead to a valid quantum map.

@@

A graphical representation for PCE maps may be introduced, being the
two qubits case what has proved to be the most useful. Consider a $N$ dimensional 
Cartesian grid, with $4^N$ places.  Each place has $N$ integer coordinates, ranging
from 0 to 3, so each place corresponds to a given $\vec \alpha$ in 
\eref{eq:N_qubits_rho}. For a given PCE, we shall fill with black 
the corresponding position \cpnote{No me pone feliz esta redaccion}
if the corresponding $\tau_{\vec \alpha}=1$. Otherwise, we leave it 
empty. Examples for $N=1$ are provided in \fref{fig:one:qubit:examples}, whereas
for two qubits, we exemplify them in \fref{fig:two:qubit:examples}. 
% Notice
% than in both cases we drop the label $\alpha$ for simplicity, as it will 
% remain unchanged throughout this text \janote{No entiendo qué estás diciendo
% aquí}.

@@

(b), (c) and (d) correspond to the identity map, \cpnote{bit flip ao algo asi}, and 

>>>>

The construction of quantum channels includes basically three ingredients:
linearity, trace preserving and complete positivity. The trace preserving
property is needed to
send convex combinations of density matrices into the convex combination of the
evolution of such density matrices, this is, let $\mcE$ such linear operation,
thus $p \rho_1 +(1-p) \rho_2 \mapsto p \mcE[\rho_1]+(1-p)\mcE[\rho_2]$. The
trace preserving property is required for the process $\mcE$ to be deterministic, this
is, $\tr\mcE[\rho]=\tr \rho =1$ is the probability that the process $\mcE$ will
happen to the state $\rho$, with $\rho \in \mcS(\hilbert)$. The complete
positivity condition is needed to preserve positive semidefiniteness and handle
the non-local character of quantum theory. This is, with positivity of the map
$\mcE$ one has $\mcE[\Delta]\geq 0 \ \ \forall \Delta \in \mcB(\hilbert)$ with
$\Delta\geq 0$\cpnote{Francois, de tu nota, la $\Delta$ ya es hermitica, pues 
decimos que es una matriz de densidad}. In the other hand, complete positivity
of $\mcE$ reads $\id_k \otimes \mcE[\omega]\geq 0$ for $k=1,2,\dots$, where
$\omega$ is the Bell state
between the system and an ancilla with Hilbert space $\hilbert_k$\flnote{No creo que lo de Bell state sea conveniente. Mejor decir algo como: 
E is totally positive if trivially extending E to the tensor product of H with an arbitrary dimensional ancilla space yields a positive operation. }. Therefore,
complete positivity ensures that every positive semidefinite operator of any
extension of the system (where the quantum channel is now a local operation) is
mapped into positive semidefinite operators, trace preserving is automatically
fulfilled for all extensions if $\mcE$ is. Thus, any density matrix of a system
where $\mcE$ acts locally, is sent to a valid density matrix, this is the
substance of complete positivity.

Quantum channels can be characterized by means of the Choi-\jami{}
isomorphism~\cite{choi,jamil}. This is, given a linear operation
$\mcE:\mcB(\hilbert)\to \mcB(\hilbert)$, the so-called Choi matrix of it,
$C_\mcE=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$, where
$\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is the
Bell state between two copies of \hilbert, is positive semi-definite if and only
if $\mcE$ is completely positive.

Quantum channels have the well-known \textit{Kraus representation} (or
operator-sum representation) which reads $\mcE[\rho]=\sum_i K_i \rho
K^\dagger{}_i$ in fact, a linear operation is completely positive if and only
if it has a Kraus representation [cite Kraus book] 
\flnote{
probablemente convenga mostrar la ecuaci'on desplegada. Asimismo se debe decir que 
$\sum_i K_i^\dagger K_i = 1$ para la trace-preserving property. 
Creo qie ``hermiticity preserving'' es en parte consecuencia de la definición, pero esto no es muy importante
}

@@

@@

Consider the linear operations that simply attenuate the Pauli components of 
a multiqubit system:
\begin{equation}
	\paulicomponents \mapsto \taus \paulicomponents.
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
\label{eq:definition:Pauli}
\end{equation}
For the operation to preserve the trace of the density matrix, we require that
$\tau_{\vec 0}=1$.  
% These operations are diagonal in the basis $\left\{
% 2^{-N/2}\sigma_{\vec \alpha} \right\}$, which follows from
% \eref{eq:sigma_property}. 
% The diagonal components are $\taus$, i.e. $\hat
% \mcE=\diag(\taus)$. Additionally, they belong to the subset of unital channels
% that can be written as convex combinations of unitary channels. \cpnote{Alguien 
% puso aca que lo probarian, pero no lo hemos discutido y quizá no le veo 
% el caso en este articulo}
\cpnote{Creo qeu vale la pena mencionar la relacion con los canales de Ruskai. 
Lo discutimos en pleno, o por lo menos cuando esten JA y FL. quizá tambien discutir
si se mencionan a los canales  \cite{Siudzinska2020}}

\cpnote{davalos, dale aca un pequeño rollo relevante quiza de una o dos
frases de la importancia de esta familia.} \ddnote{De momento no he encontrado
alguna reference, para qubits si, pero para los generales no}.

A Pauli component erasing (PCE) operation is a Pauli diagonal operation
%  A linear operation 
% $\mcE$ 
that either preserves or completely erases the Pauli components of a
density matrix. That is, 
% it transforms its Pauli components as 
\begin{equation} \label{eq:PCE_definition}
% \begin{split}
% 	\paulicomponents \mapsto \ \taus \paulicomponents,\hspace{.5cm}\\
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
	\paulicomponents \mapsto \taus \paulicomponents,\quad 
	\taus = 0,1.
% \end{split}
\end{equation}
It is worth noticing that not all PCE maps are quantum operations. For single 
qubits, if just one of the $\taus=0$, the complete positivity condition 
is not met, as discussed in \sref{sec:single:qubit}. It will be of central interest
in this article to determine which $\taus$ lead to a valid quantum map.

@@

A graphical representation for PCE maps may be introduced, being the
two qubits case what has proved to be the most useful. Consider a $N$ dimensional 
Cartesian grid, with $4^N$ places.  Each place has $N$ integer coordinates, ranging
from 0 to 3, so each place corresponds to a given $\vec \alpha$ in 
\eref{eq:N_qubits_rho}. For a given PCE, we shall fill 
the square if the corresponding $\tau_{\vec \alpha}=1$. Otherwise, we leave it 
empty. Examples for $N=1$ are provided in \fref{fig:one:qubit:examples}, whereas
for two qubits, we exemplify them in \fref{fig:two:qubit:examples}. 
% Notice
% than in both cases we drop the label $\alpha$ for simplicity, as it will 
% remain unchanged throughout this text \janote{No entiendo qué estás diciendo
% aquí}.

@@

(b), (c) and (d) correspond to the identity map, dephasing channel, and 
